---
title: "Create Seurat object out of Cell Ranger aggregation"
output: html_notehfook
---

Make directory.
```{bash}
mkdir /Volumes/Bud_SSD/20220303-early-head-fold/20190211ab-ehf-aggr-mapped
```

Download Cell Ranger aggregation.
```{bash}
scp -r schea2@hpc3.rcic.uci.edu:/share/crsp/lab/alcalof/share/dfs3-backup/lander-calof/incoming/aggregated/20190211ab-ehf-aggr-mapped/outs/filtered_feature_bc_matrix/*.gz /Volumes/Bud_SSD/20220303-early-head-fold/20190211ab-ehf-aggr-mapped
```

Load libraries.
```{r}
library(Seurat)
```

Load Cell Ranger aggregation.
```{r}
ehf.aggr <- Read10X(data.dir = "/Volumes/Bud_SSD/20220303-early-head-fold/20190211ab-ehf-aggr-mapped")
```

How many genes and cells?
```{r}
dim(ehf.aggr)
```

Generate metadata. EHF1 has 2130 cells. EHF2 has 4047 cells. 
```{r}
experiment <- vector()
experiment[1:6177] <- "20190211"
experiment <- factor(experiment)

embryo <- vector()
embryo[1:2130] <- "EHF1"
embryo[2131:6177] <- "EHF2"
embryo <- factor(embryo, levels = c("EHF1", "EHF2"))

genotype <- vector()
genotype[1:6177] <- "Flox"
genotype <- factor(genotype)

metadata <- data.frame(experiment, embryo, genotype)
rownames(metadata) <- colnames(ehf.aggr)

head(metadata)
```

Create Seurat object.
```{r}
ehf.aggr <- CreateSeuratObject(counts = ehf.aggr, project = "ehf.aggr", min.cells = 0, min.features = 0, meta.data = metadata)
```

Split object by embryo.
```{r}
embryo.aggr <- SplitObject(object = ehf.aggr, split.by = "embryo")
```

Generate pseudobulk per embryo.
```{r}
umi.embryo <- lapply(X = embryo.aggr, FUN = function(x) rowSums(x@assays$RNA@counts))
umi.embryo <- data.frame(umi.embryo[[1]], umi.embryo[[2]])
head(umi.embryo)
```

How many genes are detected in both embryos?
```{r}
d.genes = apply(umi.embryo, 1, function(row) all(row !=0 ))
sum(d.genes)
```


Filter out genes that are not detected across all embryos.
```{r}
ehf.aggr <- subset(ehf.aggr, features = rownames(ehf.aggr)[d.genes])
dim(ehf.aggr)
```

Save Seurat object.
```{r}
save(ehf.aggr, file = "/Volumes/Bud_SSD/20220303-early-head-fold/ehf-aggr-seurat.Robj")
```

