---
title: "Which cells are low quality cells?"
output: html_notebook
---

Download Seurat object.
```{bash}
#scp schea2@hpc3.rcic.uci.edu:/share/crsp/lab/alcalof/schea2/20220414-seurat/all-aggr-seurat.Robj /Volumes/all_ssd/20220414-seurat
```

Load libraries.
```{r}
library(Seurat)
library(tidyverse)
library(pracma)
library(scales)
```

Load Seurat object.
```{r}
load("/Volumes/all_ssd/20220414-seurat/all-aggr-seurat.Robj")
```

Plot histogram of number of umi's.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA)) +
  geom_histogram() +
  geom_vline(xintercept = mean(all.aggr$nCount_RNA), color = "blue") +
  geom_vline(xintercept = median(all.aggr$nCount_RNA), color = "red") +
  scale_x_continuous(trans = "log10")
```

Plot histogram of number of genes.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nFeature_RNA)) +
  geom_histogram() +
  geom_vline(xintercept = mean(all.aggr$nFeature_RNA), color = "blue") +
  geom_vline(xintercept = median(all.aggr$nFeature_RNA), color = "red") +
  scale_x_continuous(trans = "log10")
```

Calculate % mitochondrial genes.
```{r}
all.aggr[["percent.mt"]] <- PercentageFeatureSet(all.aggr, pattern = "^mt-")
head(all.aggr@meta.data)
```

Plot histogram of % mitochondrial genes.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt)) +
  geom_histogram() +
    geom_vline(xintercept = mean(all.aggr$percent.mt), color = "blue") +
  geom_vline(xintercept = median(all.aggr$percent.mt), color = "red") +
  scale_x_continuous(trans = "log10")
```
Largest portion of distribution normal. Obvious hump of cells with low % mitochondrial genes. Odd that they have low % mitochondrial genes. Usually low is better, but here, most cells have a % between 1 and 10.

% mitochondrial genes is dependent on number of genes detected. Do these correpond to cells with low numbers of genes?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA)) +
  geom_point(size = 0.1) +
      geom_vline(xintercept = mean(all.aggr$percent.mt), color = "blue") +
  geom_vline(xintercept = median(all.aggr$percent.mt), color = "red") +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
```

Calculate outliers using median absolute deviation.
```{r}
all.aggr$log.mito <- log10(all.aggr$percent.mt)
median.mito <- median(all.aggr$log.mito)
all.aggr$d.mito <- abs(all.aggr$log.mito - median.mito)
md.mito <- mad(all.aggr$log.mito)
all.aggr$mito.out <- all.aggr$d.mito > md.mito*3
sum(all.aggr$mito.out)
```

```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA, color = mito.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")

```

What does histogram of % mito now look?
```{r}
mito.hist <- ggplot() +
  geom_histogram(data = all.aggr@meta.data, mapping = aes(x = percent.mt, fill = mito.out, color = mito.out), alpha = 0.5, size = 0.33) +
  scale_x_continuous(trans = "log10", labels = scales::percent_format(scale = 1)) +
  scale_fill_discrete(labels = c("<=3 MAD", "> 3 MAD")) +
  scale_color_discrete(labels = c("<=3 MAD", "> 3 MAD")) +
  labs(x = "% Mitochondrial Genes", y = "Number of Cells", fill = "", color = "", tag = "A") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.key.size = unit(7, "pt"), legend.position = "top", plot.tag = element_text(size = 9, face = "bold"))

mito.hist

ggsave(plot = mito.hist, filename = "mito-hist-plot-low-quality-cells.png", device = "png", units = "in", width = 2.16, height = 3)

```

Genes?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nFeature_RNA, color = mito.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```

UMIs?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, color = mito.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```

What does vln plot mito now look?
```{r}
mito.plot <- ggplot() +
  geom_jitter(data = all.aggr@meta.data, mapping = aes(y = percent.mt, x = mito.out, group = mito.out), stroke = 0, size = 0.25) +
  geom_violin(data = all.aggr@meta.data, mapping = aes(y = percent.mt, x = mito.out, color = mito.out), alpha = 0) +
  scale_x_discrete(labels = c("<= 3MAD", "> 3MAD")) +
  scale_y_continuous(trans = "log10", labels = scales::percent_format(scale = 1)) +
  labs(x = "", y = "% Mitochondrial Genes", fill = "", color = "", tag = "A") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none", plot.tag = element_text(size = 9, face = "bold"))

mito.plot

ggsave(plot = mito.plot, filename = "mito-vln-plot-low-quality-cells.png", device = "png", units = "in", width = 2.16, height = 4.5)
```

Can we find outliers among genes after already filtering out outliers from mito?
```{r}
all.aggr$log.gene <- log10(all.aggr$nFeature_RNA)
median.gene <- median(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% pull(log.gene))
all.aggr$d.gene <- abs(all.aggr$log.gene - median.gene)
md.gene <- mad(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% pull(log.gene))
all.aggr$gene.out <- all.aggr$d.gene > md.gene*3
sum(all.aggr$gene.out)
```

Where do these cells lie along histogram of number of genes?
```{r}
gene.hist <- ggplot() +
  geom_histogram(data = all.aggr@meta.data, mapping = aes(x = nFeature_RNA, fill = gene.out, color = gene.out), alpha = 0.5, size = 0.33) +
  scale_x_continuous(trans = "log10", breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_fill_discrete(labels = c("<=3 MAD", "> 3 MAD")) +
  scale_color_discrete(labels = c("<=3 MAD", "> 3 MAD")) +
  labs(x = "Number of Genes", y = "Number of Cells", fill = "", color = "", tag = "B") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.key.size = unit(7, "pt"), legend.position = "top", plot.tag = element_text(size = 9, face = "bold"))

gene.hist 

ggsave(plot = gene.hist, filename = "gene-hist-plot-low-quality-cells.png", device = "png", units = "in", width = 2.16, height = 3)


```

Along number of umis?
```{r}
ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE), mapping = aes(x = nCount_RNA, color = gene.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```

What does vln plot genes now look?
```{r}
gene.plot <- ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE)) +
  geom_jitter(mapping = aes(y = nFeature_RNA, x = gene.out, group = gene.out), stroke = 0, size = 0.25) +
  geom_violin(mapping = aes(y = nFeature_RNA, x = gene.out, color = gene.out), alpha = 0) +
  scale_x_discrete(labels = c("<= 3MAD", "> 3MAD")) +
  scale_y_continuous(trans = "log10", breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  labs(x = "", y = "Number of Genes", fill = "", color = "", tag = "B") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none", , plot.tag = element_text(size = 9, face = "bold"))

gene.plot

ggsave(plot = gene.plot, filename = "gene-vln-plot-low-quality-cells.png", device = "png", units = "in", width = 2.16, height = 4.5)
```

Can we find outliers among umis after already filtering out mito and genes?
```{r}
all.aggr$log.umi <- log10(all.aggr$nCount_RNA)
median.umi <- median(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE) %>% pull(log.umi))
all.aggr$d.umi <- abs(all.aggr$log.umi - median.umi)
md.umi <- mad(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE) %>% pull(log.umi))
all.aggr$umi.out <- all.aggr$d.umi > md.umi*3
sum(all.aggr$umi.out)
```

Where do these lie along umis?
```{r}
umi.hist <- ggplot() +
  geom_histogram(data = all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE), mapping = aes(x = nCount_RNA, color = umi.out, fill = umi.out), alpha = 0.5, size = 0.33) +
  scale_x_continuous(trans = "log10", breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  scale_fill_discrete(labels = c("<=3 MAD", "> 3 MAD")) +
  scale_color_discrete(labels = c("<=3 MAD", "> 3 MAD")) +
  labs(x = "Number of UMIs", y = "Number of Cells", fill = "", color = "", tag = "C") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.key.size = unit(7, "pt"), legend.position = "top", plot.tag = element_text(size = 9, face = "bold"))

umi.hist

ggsave(plot = umi.hist, filename = "umi-hist-plot-low-quality-cells.png", device = "png", units = "in", width = 2.16, height = 3)
```

Plot dots.
```{r}
ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE), mapping = aes(x = nCount_RNA, y = 1, color = umi.out)) +
  geom_point(position = position_jitter(), size = 0.1) +
  scale_x_continuous(trans = "log10")
```
So some upper and some lower range cells. 

What does vln plot genes now look?
```{r}
umi.plot <- ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE)) +
  geom_jitter(mapping = aes(y = nCount_RNA, x = umi.out, group = umi.out), stroke = 0, size = 0.1) +
  geom_violin(mapping = aes(y = nCount_RNA, x = umi.out, color = umi.out), alpha = 0) +
  scale_x_discrete(labels = c("<= 3MAD", "> 3MAD")) +
  scale_y_continuous(trans = "log10", breaks=trans_breaks('log10', function(x) 10^x), labels=trans_format('log10', math_format(10^.x))) +
  labs(x = "", y = "Number of UMIs", fill = "", color = "", tag = "C") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.position = "none", plot.tag = element_text(size = 9, face = "bold"))

umi.plot

ggsave(plot = umi.plot, filename = "umi-vln-plot-low-quality-cells.png", device = "png", units = "in", width = 2.16, height = 4.5)
```

So a total of 768 for umi, 65 for gene, and 6681 for mito totaling 7457 cells. There is some overlap.

But there may be some overlap. Which cells are out for any one of these reasons?
```{r}
all.aggr$all.out <- (all.aggr$mito.out == TRUE | all.aggr$gene.out == TRUE | all.aggr$umi.out == TRUE)
sum(all.aggr$all.out)
```
So those are the cells to filter out. 

Where are these cells along mito and genes?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
```

```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") + 
  scale_y_continuous(trans = "log10") + 
  facet_wrap(~label.embryo)
```

How about along gene and umi?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10") +
  facet_wrap(~all.out)
```

```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") +
  scale_y_continuous(trans = "log10")
  facet_wrap(~label.embryo)
```

Let's remove those cells.
```{r}
all.out <- all.aggr$all.out
save(all.out, file = "/Volumes/all_ssd/20220414-seurat/all-out-all-aggr.Robj")
```

```{bash}
scp /Volumes/all_ssd/20220414-seurat/all-out-all-aggr.Robj schea2@hpc3.rcic.uci.edu:/share/crsp/lab/alcalof/schea2/20220414-seurat
```

```{r}
high.cells <- all.aggr@meta.data %>% filter(all.out == FALSE)
head(high.cells)
```

How many cells per embryo?
```{r}
table(high.cells$label.embryo)
```

Median UMIs per cell per embryo?
```{r}
high.cells %>% group_by(label.embryo) %>% summarise(median(nCount_RNA))
```

Median genes per cell per embryo?
```{r}
high.cells %>% group_by(label.embryo) %>% summarise(median(nFeature_RNA))
```

Median UMIs per cell per stage per genotype?
```{r}
high.cells %>% group_by(stage, genotype) %>% summarise(median(nCount_RNA))
```

Median genes per cell per stage per genotype?
```{r}
high.cells %>% group_by(stage, genotype) %>% summarise(median(nFeature_RNA))
```

Median UMIs per cell per stage per genotype?
```{r}
high.cells %>% group_by(stage) %>% summarise(median(nCount_RNA))
```

Median genes per cell per stage per genotype?
```{r}
high.cells %>% group_by(stage) %>% summarise(median(nFeature_RNA))
```

Median UMIs per cell per stage per genotype?
```{r}
high.cells %>% group_by(genotype) %>% summarise(median(nCount_RNA))
```

Median genes per cell per stage per genotype?
```{r}
high.cells %>% group_by(genotype) %>% summarise(median(nFeature_RNA))
```

Median UMIs per cell?
```{r}
median(high.cells$nCount_RNA)
```

Median genes per cell?
```{r}
median(high.cells$nFeature_RNA)
```



