---
title: "Confirm cluster iterations were performed at a low enough resolution"
output: html_notebook
---

Download cluster iterations.
```{bash}
scp -r schea2@hpc3.rcic.uci.edu:/share/crsp/lab/alcalof/schea2/20220414-seurat/all-aggr-cluster-iters /Volumes/all_ssd/20220414-seurat
```

Load libraries.
```{r}
library(tidyverse)
theme_set(theme_classic(base_size = 7))
library(igraph)
library(ggraph)
```

At each resolution, calculate shannon entropy per cluster.
```{r}
max.res <- 0.2

#generate empty list to store cluster identities for each cell
cluster.cell.list <- list()

#open cluster-per-cell files and store them in list above
for (i in seq(from = 0, to = max.res, by = 0.0001)) {
  file.name <- paste0("/Volumes/all_ssd/20220414-seurat/all-aggr-cluster-iters/cluster-per-cell-all-aggr-res-", sprintf(fmt = "%.4f", i), ".Robj")
  load(file.name)
  cluster.cell.list[[sprintf(fmt = "%.4f", i)]] <- as.numeric(all.aggr.active.ident)
}

#generate matrix of cluster ids per resolution across columns
cluster.cell.matrix <- bind_cols(cluster.cell.list)

#generate tidy data of node ids and attributes
tidy.node <- as.data.frame(unlist(apply(cluster.cell.matrix, 2, table)))
colnames(tidy.node) <- "node.size"
tidy.node$node <- row.names(tidy.node)
tidy.node <- separate(tidy.node, col = node, sep = "\\.(?!.*\\.)" , into = c("resolution", "cluster"), remove = FALSE)

res <- sprintf(fmt = "%.4f", seq(from = 0, to = max.res, by = 0.0001))

tidy.node$resolution <- factor(tidy.node$resolution, levels = res)
tidy.node$cluster <- factor(tidy.node$cluster, levels = 1:max(tidy.node$cluster))

#shannon value of resolution 1 must be zero by definition
shannon.per.cluster <- vector()
shannon.per.cluster <- c(0)
names(shannon.per.cluster) <- "1"
shannon.per.cluster.per.res <- list()
shannon.per.cluster.per.res[[res[1]]] <- shannon.per.cluster

#calculate diversity score for each resolutoin beginning at 3
for (i in 1:(length(res)-1)) {
  res.subset.cluster.cell.matrix <- as.data.frame(cluster.cell.matrix[, c(i, i+1)])
  colnames(res.subset.cluster.cell.matrix) <- c("res.low", "res.high")
  table.res.subset.cluster.cell.matrix <- table(res.subset.cluster.cell.matrix[, "res.high"])
  
  shannon.per.cluster <- vector()
  
  for (t in names(table.res.subset.cluster.cell.matrix)) {
    clust.subset.cluster.cell.matrix <- filter(res.subset.cluster.cell.matrix,  res.high == t)
    clust.subset.cluster.cell.matrix$res.low <- factor(clust.subset.cluster.cell.matrix$res.low)
    table.clust.subset.cluster.cell.matrix <- table(clust.subset.cluster.cell.matrix$res.low)
    shannon.per.cluster[t] <- sum(-(table.clust.subset.cluster.cell.matrix/table.res.subset.cluster.cell.matrix[t])*log(table.clust.subset.cluster.cell.matrix/table.res.subset.cluster.cell.matrix[t]))
  }
  
  shannon.per.cluster.per.res[[res[i+1]]] <- shannon.per.cluster
}
tidy.node$shannon <- unlist(shannon.per.cluster.per.res)
tidy.node <- tidy.node[, c("node", "resolution", "cluster", "node.size", "shannon")]
row.names(tidy.node) <- NULL
```

Calculate total shannon entropy per resolution.
```{r}
total.shannon <- tidy.node %>% group_by(resolution) %>% summarise(shannon = sum(shannon))
head(total.shannon)
```

Add number of clusters per resolution.
```{r}
total.shannon$clusters <- tidy.node %>% group_by(resolution) %>% summarise(clusters = length(cluster)) %>% pull(clusters)
head(total.shannon)
```

Grab only those resolutions with a total shannon entropy of zero.
```{r}
zero.shannon <- total.shannon %>% filter(shannon == 0)
head(zero.shannon)
```

For each number of clusters, how many resolutions had a total shannon entropy of zero?
```{r}
table(zero.shannon$clusters)
```
Since we consecutively clustered cells at every number of clusters, we used a low enough resolution. 

Rank number of clusters by number of resolutions from high to low.
```{r}
sort(table(zero.shannon$clusters), decreasing = TRUE)
```

Heat map can only fit 79 genes. For 10 genes each cluster, the max clusters you can show is 7. For 5 genes each cluster, the max clusters you can show is 15. So we can't cluster at 20 or 17 anyway.

How do we find the best resolution for each number of clusters?

Grab all resolutions at 2 clusters.
```{r}
res.per.clust <- as.numeric(as.character(zero.shannon %>% filter(clusters == 3) %>% pull(resolution)))
res.per.clust
```

Calculate segments. 
```{r}
seg.per.clust <- c(0, which(as.integer(diff(x = res.per.clust / 0.0001)) > 1), length(res.per.clust))
seg.per.clust
```

Calculate distance of each segment.
```{r}
dis.per.seg <- diff(x = seg.per.clust)
dis.per.seg
```

Grab best resoluton at each cluster.
```{r}
best.res.per.clust <- res.per.clust[seg.per.clust[which.max(dis.per.seg) + 1]]
best.res.per.clust
```

```{r}
best.res.per.clust <- vector()

for (i in 1:42) {
  res.per.clust <- as.numeric(as.character(zero.shannon %>% filter(clusters == i) %>% pull(resolution)))
  seg.per.clust <- c(0, which(as.integer(diff(x = res.per.clust / 0.0001)) > 1), length(res.per.clust))
  dis.per.seg <- diff(x = seg.per.clust)
  best.res.per.clust <- res.per.clust[seg.per.clust[which.max(dis.per.seg) + 1]]
  print(i)
  print(best.res.per.clust)
}

```






