---
title: "Find variable genes"
output: html_notebook
---

Download variable genes.
```{bash}
scp schea2@hpc3.rcic.uci.edu:/share/crsp/lab/alcalof/schea2/20220414-seurat/var-genes-flox-aggr-sctrans.Robj .
```

Load variable genes.
```{r}
load("var-genes-flox-aggr-sctrans.Robj")
```

How many variable genes per embryo?
```{r}
print("number of variable geness per embryo")
lapply(var.genes, FUN = length)
```



Find unique genes among all embryos.
```{r}
uniq.genes <- unique(c(var.genes[[1]],
var.genes[[2]],
var.genes[[3]],
var.genes[[4]],
var.genes[[5]],
var.genes[[6]],
var.genes[[7]],
var.genes[[8]],
var.genes[[9]],
var.genes[[10]],
var.genes[[11]],
var.genes[[12]],
var.genes[[13]]
))

print("number of unique genes among all embryos")
length(uniq.genes)
```

Count how many times each gene is counted as variable among all embryos. Rank from most common to least common. 
```{r}
print("counting how many times each gene occurs")
count.gene <- list()
for (e in uniq.genes) {
    for(r in names(var.genes)) {
        count.gene[[e]][[r]]<- e %in% var.genes[[r]]
    }
}

print("ranking unique variable genes")
count.rank <- lapply(X = count.gene, FUN = unlist)
count.rank <- lapply(X = count.rank, FUN = sum)
count.rank <- sort(unlist(count.rank), decreasing = TRUE)

print("table of rankings")
print(table(count.rank))
```

Find unique genes among late bud embryos.
```{r}
lb.uniq.genes <- unique(c(var.genes[[1]],
var.genes[[2]],
var.genes[[3]],
var.genes[[4]],
var.genes[[5]]
))

print("number of unique genes among late bud embryos")
length(lb.uniq.genes)
```

Count how many times each gene is counted as variable among late bud embryos. Rank from most common to least common. 
```{r}
print("counting how many times each gene occurs")
lb.count.gene <- list()
for (e in lb.uniq.genes) {
    for(r in names(var.genes)[1:5]) {
        lb.count.gene[[e]][[r]]<- e %in% var.genes[[r]]
    }
}

print("ranking unique variable genes")
lb.count.rank <- lapply(X = lb.count.gene, FUN = unlist)
lb.count.rank <- lapply(X = lb.count.rank, FUN = sum)
lb.count.rank <- sort(unlist(lb.count.rank), decreasing = TRUE)

print("table of rankings")
print(table(lb.count.rank))
```

Grab variable genes common to all late bud embryos.
```{r}
lb.var.genes <- names(lb.count.rank[lb.count.rank >= 5])
head(lb.var.genes)
```

Find unique genes among cardiac crescent embryos.
```{r}
cc.uniq.genes <- unique(c(var.genes[[6]],
var.genes[[7]],
var.genes[[8]],
var.genes[[9]],
var.genes[[10]],
var.genes[[11]],
var.genes[[12]],
var.genes[[13]]
))

print("number of unique genes among cardiac crescent embryos")
length(cc.uniq.genes)
```

Count how many times each gene is counted as variable among cardiac crescent embryos. Rank from most common to least common. 
```{r}
print("counting how many times each gene occurs")
cc.count.gene <- list()
for (e in cc.uniq.genes) {
    for(r in names(var.genes)[6:13]) {
        cc.count.gene[[e]][[r]]<- e %in% var.genes[[r]]
    }
}

print("ranking unique variable genes")
cc.count.rank <- lapply(X = cc.count.gene, FUN = unlist)
cc.count.rank <- lapply(X = cc.count.rank, FUN = sum)
cc.count.rank <- sort(unlist(cc.count.rank), decreasing = TRUE)

print("table of rankings")
print(table(cc.count.rank))
```

Grab variable genes common to all cardiac crescent embryos.
```{r}
cc.var.genes <- names(cc.count.rank[cc.count.rank >= 8])
head(cc.var.genes)
```

Grab variable genes common to both late bud and cardiac crescent embryos.
```{r}
all.var.genes <- unique(c(lb.var.genes, cc.var.genes))
head(all.var.genes)
length(all.var.genes)
```



















