---
title: "Which cells are low quality cells?"
output: html_notebook
---

Load libraries.
```{r}
library(Seurat)
library(tidyverse)
library(pracma)
```

Load seurat object.
```{r}
load("/Volumes/Bud_SSD/20210505-seurat/all-aggr-seurat.Robj")
```

Explore meta data.
```{r}
head(all.aggr@meta.data)
```

Plot histogram of number of umi's.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA)) +
  geom_histogram() +
  geom_vline(xintercept = mean(all.aggr$nCount_RNA), color = "blue") +
  geom_vline(xintercept = median(all.aggr$nCount_RNA), color = "red") +
  scale_x_continuous(trans = "log10")
```
Looks normal except for bump on left.

Plot histogram of number of umi's. Color by embryo.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, color = embryo.id)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10") + 
  facet_wrap(~embryo.id)
```
Bump on left due to ATU6-45 and BBH6-59.

What are the genotypes of these two embryos? Which experiments did they come from?
```{r}
print("ATU6-45")
(all.aggr@meta.data %>% filter(embryo.id == "ATU6-45") %>% pull(genotype))[1]
(all.aggr@meta.data %>% filter(embryo.id == "ATU6-45") %>% pull(experiment))[1]

print("AYR4-22")
(all.aggr@meta.data %>% filter(embryo.id == "AYR4-22") %>% pull(genotype))[1]
(all.aggr@meta.data %>% filter(embryo.id == "AYR4-22") %>% pull(experiment))[1]
```
Both mutant. But from two different experiments.

Plot histogram of number of genes.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nFeature_RNA)) +
    geom_histogram() +
    geom_vline(xintercept = mean(all.aggr$nFeature_RNA), color = "blue") +
  geom_vline(xintercept = median(all.aggr$nFeature_RNA), color = "red")

```
Looks similar to histogram of number of umi's. More normal. Except for bump on left again.

Plot histogram of number of genes. Color by embryo.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nFeature_RNA, color = embryo.id)) +
  geom_histogram() +
  facet_wrap(~embryo.id)
```
Bump is due to same two embryos earlier. ATU6-45 and AYR4-22. 

Calculate % mitochondrial genes.
```{r}
all.aggr[["percent.mt"]] <- PercentageFeatureSet(all.aggr, pattern = "^mt-")
head(all.aggr@meta.data)
```

Plot histogram of % mitochondrial genes.
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt)) +
  geom_histogram() +
    geom_vline(xintercept = mean(all.aggr$percent.mt), color = "blue") +
  geom_vline(xintercept = median(all.aggr$percent.mt), color = "red") +
  scale_x_continuous(trans = "log10")
```
Largest portion of distribution normal. Obvious hump of cells with low % mitochondrial genes. Odd that they have low % mitochondrial genes. Usually low is better, but here, most cells have a % between 1 and 10.

Are these cells associated with one embryo?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, color = embryo.id)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10") +
  facet_wrap(~embryo.id, scales = "free_y")
```
Left bump due to more than one embryo. Pervasive feature.

% mitochondrial genes is dependent on number of genes detected. Do these correpond to cells with low numbers of genes?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA)) +
  geom_point(size = 0.1) +
      geom_vline(xintercept = mean(all.aggr$percent.mt), color = "blue") +
  geom_vline(xintercept = median(all.aggr$percent.mt), color = "red") +
  scale_x_continuous(trans = "log10")
```
So the cells with low mito tend to have lower number of genes. mito seems to distribute normally along number of genes. But if we only filter these out, it doesn't account for upper range. 

Calculate outliers using median absolute deviation.
```{r}
all.aggr$log.mito <- log10(all.aggr$percent.mt)
median.mito <- median(all.aggr$log.mito)
all.aggr$d.mito <- abs(all.aggr$log.mito - median.mito)
md.mito <- mad(all.aggr$log.mito)
all.aggr$mito.out <- all.aggr$d.mito > md.mito*3
sum(all.aggr$mito.out)
```

```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA, color = mito.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10")
```

What does histogram of % mito now look?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, color = mito.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```

Genes?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nFeature_RNA, color = mito.out)) +
  geom_histogram()
```

UMIs?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, color = mito.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```

Can we find outliers among genes after already filtering out outliers from mito?
```{r}
median.gene <- median(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% pull(nFeature_RNA))
all.aggr$d.gene <- abs(all.aggr$nFeature_RNA - median.gene)
md.gene <- mad(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% pull(nFeature_RNA))
all.aggr$gene.out <- all.aggr$d.gene > md.gene*3
sum(all.aggr$gene.out)
```
So 221 cells can be filtered out after already filtering out outliers from mito.

Where do these cells lie along histogram of number of genes?
```{r}
ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE), mapping = aes(x = nFeature_RNA, color = gene.out)) +
  geom_histogram()
```
Captured those in upper range.

Along number of umis?
```{r}
ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE), mapping = aes(x = nCount_RNA, color = gene.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```
Upper again.

Can we find outliers among umis after already filtering out mito and genes?
```{r}
all.aggr$log.umi <- log10(all.aggr$nCount_RNA)
median.umi <- median(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE) %>% pull(log.umi))
all.aggr$d.umi <- abs(all.aggr$log.umi - median.umi)
md.umi <- mad(all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE) %>% pull(log.umi))
all.aggr$umi.out <- all.aggr$d.umi > md.umi*3
sum(all.aggr$umi.out)
```
So another 197 cells after already filering out outliers from mito and genes.

Where do these lie along umis?
```{r}
ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE), mapping = aes(x = nCount_RNA, color = umi.out)) +
  geom_histogram() +
  scale_x_continuous(trans = "log10")
```
Hard to see where these are.

Plot dots.
```{r}
ggplot(data = all.aggr@meta.data %>% filter(mito.out == FALSE) %>% filter(gene.out == FALSE), mapping = aes(x = nCount_RNA, y = 1, color = umi.out)) +
  geom_point(position = position_jitter(), size = 0.1) +
  scale_x_continuous(trans = "log10")
```
So some upper and some lower range cells. 

So a total of 197 for umi, 221 for gene, and 2256 for mito totaling 2511 cells. There is some overlap.

But there may be some overlap. Which cells are out for any one of these reasons?
```{r}
all.aggr$all.out <- (all.aggr$mito.out == TRUE | all.aggr$gene.out == TRUE | all.aggr$umi.out == TRUE)
sum(all.aggr$all.out)
```
So those are the cells to filter out. 

Where are these cells along mito and genes?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10")
```

```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = percent.mt, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") + 
  facet_wrap(~embryo.id)
```

How about along gene and umi?
```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") +
  facet_wrap(~all.out)
```

```{r}
ggplot(data = all.aggr@meta.data, mapping = aes(x = nCount_RNA, y = nFeature_RNA, color = all.out)) +
  geom_point(size = 0.1) +
  scale_x_continuous(trans = "log10") +
  facet_wrap(~embryo.id)
```

Let's remove those cells.
```{r}
all.out <- all.aggr$all.out
save(all.out, file = "/Volumes/Bud_SSD/20210505-seurat/all-out-all-aggr.Robj")
```

```{bash}
scp /Volumes/Bud_SSD/20210505-seurat/all-out-all-aggr.Robj schea2@hpc3.rcic.uci.edu:/dfs6/pub/schea2/20210505-seurat
```

