---
title: "Do FIN embryos express Nipbl at lower levels than Flox embryos?"
output: html_notebook
---

Load required libraries.
```{r}
library(tidyverse)
library(Seurat)
library(ggsignif)
library(Rmisc)
library(scales)
```

Load Seurat object.
```{r}
load("/Volumes/Bud_SSD/20210505-seurat/all-aggr-sctrans.Robj")
```

How many cells?
```{r}
dim(all.aggr)
```

How are the cells grouped?
```{r}
head(all.aggr@active.ident)
```

Group the cells by embryo.
```{r}
Idents(all.aggr) <- all.aggr$embryo.id
```

Split Flox cells by embryo.
```{r}
embryo.aggr <- SplitObject(all.aggr, split.by = "ident")
```

Grab embryo data.
```{r}
umi.embryo <- lapply(X = embryo.aggr, FUN = function(x) rowSums(x@assays$SCT_ALL@counts))
```

Organize as data frame.
```{r}
umi.data <- data.frame(umi.embryo[[1]], umi.embryo[[2]], umi.embryo[[3]], umi.embryo[[4]], umi.embryo[[5]], umi.embryo[[6]], umi.embryo[[7]], umi.embryo[[8]], umi.embryo[[9]], umi.embryo[[10]], umi.embryo[[11]])
head(umi.data)
```

Rename columns.
```{r}
colnames(umi.data) <- names(umi.embryo)
head(umi.data)
```


Is the total depth per embryo the same?
```{r}
plot(colSums(umi.data))
```
It is not. 

Calculate normalization factor using minimum. 
```{r}
total.embryo <- colSums(umi.data)
norm.factor <- total.embryo / min(total.embryo)
norm.factor
```

Divide each gene by scaling factor. 
```{r}
norm.umi <- as.data.frame(apply(X = umi.data, MARGIN = 1, FUN = function(x) x/norm.factor))
head(norm.umi)
```

Add genotype.
```{r}
norm.umi$genotype <- c("Flox", "Flox", "Flox", "Flox", "Flox", "FIN", "FIN", "FIN", "FIN", "FIN", "FIN")
norm.umi$genotype <- factor(norm.umi$genotype, levels = c("Flox", "FIN"))
```

Let's calculate the mean and confidence interval.
```{r}
error.nipbl <- summarySE(data = norm.umi, measurevar = "Nipbl", groupvars = "genotype")
error.nipbl
```

Let's test the difference in Nipbl expression.
```{r}
t.test(formula = log10(norm.umi$Nipbl) ~ norm.umi$genotype, alternative = "two.sided", paired = FALSE, conf.level = 0.95)
```

```{r}
nipbl.plot <- ggplot() +
  geom_jitter(data = norm.umi, mapping = aes(x = genotype, y = Nipbl, group = genotype), width = 0.1) +
  geom_point(data = error.nipbl, mapping = aes(x = genotype, y = Nipbl, color = genotype, group = genotype), position = position_nudge(x = 0.15)) +
  geom_errorbar(data = error.nipbl, mapping = aes(x = genotype, ymin = Nipbl-ci, ymax = Nipbl+ci, color = genotype, group = genotype), width = 0.05, position = position_nudge(x = 0.15)) +
  scale_color_manual(values = c("blue", "red"), labels = c("Nipbl Flox/+", "Nipbl FIN/+")) + 
  geom_signif(data = norm.umi, mapping = aes(x = genotype, y = Nipbl, group = genotype), xmin = 1.15, xmax = 2.15, y_position = 3.35, annotation = "**", size = 0.33, textsize = 3.5) +
  scale_x_discrete(labels = c("Nipbl Flox/+\nN = 5", "Nipbl FIN/+\nN = 6")) +
  scale_y_log10(breaks=trans_breaks('log10', function(x) 10^x),
                     labels=trans_format('log10', math_format(10^.x))) +
  labs(title = "Nipbl", x = "Genotype", y = "Norm UMI Per Embryo", color = "95% CI Mean") +
  theme_classic(base_size = 7) +
  theme(plot.title = element_text(hjust = 0.5), legend.key.height = unit(7, "pt"))
nipbl.plot
```

```{r}
ggsave(plot = nipbl.plot, filename = "/Volumes/Bud_SSD/20210505-seurat/nipbl-umi-between-genotype-all-aggr.png", device = "png", units = "in", width = 6.5, height = 4.5)
```
